### 2018-07-01 paper/4-build.tex
Work on S4

An alternative way to determine if a key is dirty is to record what state the
values/hashes of dependencies were used at last time, and if something has
changed, the key is dirty and must be recomputed -- in essence a \emph{trace}
which we can use to \emph{verify} existing values. We can describe a trace as:

We assume that \hs{Hash}~\hs{v} is a small constant size, constructed from hashing the
underlying \hs{v} rather than storing it directly. Checking a trace requires
ensuring all the dependencies are up to date (using whatever ordering strategy
as per \S\ref{sec-dependency-orderings}), then comparing if the dependencies are
same as the current value and that the result is the same.

A build system that uses verifying traces needs to persistently maintain a set
of traces. After computing a fresh value we add its \hs{Trace}~\hs{k}~\hs{v} to
the set. Therefore the information stored by a build system that verifies
traces can be modelled as a list or set~--~we chose \hs{[Trace}~\hs{k}~\hs{v]}
for simplicity. In practice, different build systems can optimise the data
structures used by traces for their specific use cases, which we discuss
in~\S\ref{sec-smart-traces}.

@@

A verifying trace allows us to mark a key dirty and rebuild it. Extending that information we can store a \emph{constructive} trace which is the trace plus the actual result. Once we are storing the complete result it makes sense to record many constructive traces per key, and to share them with other users, providing cloud-build functionality. We can represent that by changing \hs{Trace} to:

\begin{minted}{haskell}
data Trace k v = Trace
    { key     :: k
    , depends :: [(k, Hash v)]
    , result  :: v }
\end{minted}

The change from~\S\ref{sec-verifying-traces} is that instead of storing the \hs{Hash}~\hs{v} for \hs{result}, we now store the full value \hs{v}. Checking a trace follows the same pattern as before, but if the result is the only thing that is different we can simply use the result \emph{without} recomputing it.

>>>>

An alternative way to determine if a key is dirty is to record what state the
values/hashes of dependencies were used at last time, and if something has
changed, the key is dirty and must be recomputed -- in essence a \emph{trace}
which we can use to \emph{verify} existing values. For traces, there are two essential
operations -- adding a new value to the trace store, and using the traces to determine if something needs
rebuilding. Assuming a store of verifying traces \hs{VT}~\hs{k}~\hs{v}, the operations are:

\begin{minted}{haskell}
recordVT :: k -> Hash v -> [(k, Hash v)] -> VT k v -> VT k v
verifyVT :: (Monad m, Eq k, Eq v)
         => k -> Hash v -> (k -> m (Hash v)) -> VT k v -> m Bool
\end{minted}

We assume that \hs{Hash}~\hs{v} is a small constant size, constructed from hashing the
underlying \hs{v} rather than storing it directly. After successfully building a key, we call \hs{recordVT} passing the key, the hash of its value, and the list of hashes and dependencies. Combined with the existing set of verifying traces, that provides a new set of verifying traces.

More interestingly, to \emph{verify} whether a key needs rebuilding we supply the key and the hash of the current result in the store, along with a function for obtaining the post-build result of any key (using whatever ordering strategy
as per \S\ref{sec-dependency-orderings}), and the existing \hs{VT} information. The result will be a \hs{Bool} where \hs{True} indicates that the trace could be verified, or \hs{False} to indicate the key should be rebuilt.

One potential implementation would be to record all arguments passed to \hs{recordVT}, and verify by simply checking if any argument matches the information passed by \hs{verifyVT} (we discuss smarter implementations in~\S\ref{sec-smart-traces}).

@@

A verifying trace allows us to mark a key dirty and rebuild it. Extending that information we can store a \emph{constructive} trace which is the trace plus the actual result. Once we are storing the complete result it makes sense to record many constructive traces per key, and to share them with other users, providing cloud-build functionality. We can represent this additional information by providing the operations:

\begin{minted}{haskell}
recordCT :: k -> v -> [(k, Hash v)] -> CT k v -> CT k v
constructCT :: (Monad m, Eq k, Eq v) => k -> (k -> m (Hash v)) -> CT k v -> m [v]
\end{minted}

The function \hs{recordCT} looks like \hs{recordVT}, but instead of just passing the hash of the result, we require the whole result. The \hs{verifyVT} has been replaced with \hs{constructCT}, which instead of taking the hash of the value, instead returns a list of suitable values. If the current value in the store matches one of the possible values, the build can skip this rule. If the resulting list is empty, the build must rerun. However, if the current value does not match the store, and there is a possible value, we can use any value from the constructive list \emph{without} doing any work to compute it.
